---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

First we set the locale to english
```{r, results='hide'}
Sys.setlocale("LC_TIME", "C")
```

then we load and cache the data

```{r, cache=TRUE}
data <- read.csv("./activity.csv")
```

and finally we preprocess the data

```{r}
data$realDate <- as.Date(data[, 'date'], format='%Y-%m-%d')
pdata <- subset(data, !is.na(steps))
```

## What is mean total number of steps taken per day?

We plot an histogram of the total number of steps taken each day
```{r, fig.width=4, fig.height=3, message=FALSE, fig.align='center', }
stepsPerDay <- aggregate(pdata$steps, by=list(date=pdata$date), FUN=sum)
hist(stepsPerDay$x, xlab="Steps", main="")
```

We calculate the mean and the median 
```{r, results='hide' }
mean1 <- format(mean(stepsPerDay$x, na.rm=T), nsmall=2)
median1 <- format(median(stepsPerDay$x, na.rm=T), nsmall=2)
```

The mean number of steps taken per day is around `r mean1`.

The median is `r median1`.


## What is the average daily activity pattern?
```{r, fig.width=4, fig.height=3, message=FALSE, fig.align='center', }
stepsPerInterval <- aggregate(pdata$steps, by=list(interval=pdata$interval), FUN=mean)
plot(stepsPerInterval$interval, stepsPerInterval$x, type="l", xlab="Interval", ylab="Steps", main="")
```


On average the interval with the maximum number of steps is the number `r stepsPerInterval[which.max(stepsPerInterval$x), 'interval']`.

## Imputing missing values

```{r}
isnas <- sum(is.na(data$steps))
```

There are `r isnas` missing values (where the steps values is NA).

We can replace every NA value with the equivalent interval average.
```{r}
data$allSteps <- ifelse(is.na(data$steps), stepsPerInterval[match(data$interval, stepsPerInterval$interval), 'x'], data$steps) 
```

We plot an histogram of the total number of steps taken each day
```{r, fig.width=4, fig.height=3, message=FALSE, fig.align='center', }
allStepsPerDay <- aggregate(data$allSteps, by=list(date=data$date), FUN=sum)
hist(allStepsPerDay$x, xlab="Steps", main="")
```

We calculate the mean and the median 
```{r, results='hide' }
mean2 <- format(mean(allStepsPerDay$x, na.rm=T), nsmall=2)
median2 <- format(median(allStepsPerDay$x, na.rm=T), nsmall=2)
```

The mean number of steps taken per day is around `r mean2`.

The median is `r median2`.

Replacing the missing values with the mean of the equivalent interval does not alter the mean and the median of steps taken per day.

## Are there differences in activity patterns between weekdays and weekends?

We split the data beased on the fact that the date is a weekday or a day of the weekend.
```{r, results='hide'}
data$weekday <- ifelse(weekdays(data$realDate, abbreviate=T) %in% c('Sat', 'Sun'), 'weekend', 'weekday')
weekdayData <- subset(data, weekday=='weekday')
weekendData <- subset(data, weekday=='weekend')
```
And we plot the data
```{r, fig.width=6, fig.height=6, message=FALSE, fig.align='center', }
stepsPerWeekdayInterval <- aggregate(weekdayData$allSteps, by=list(interval=weekdayData$interval), FUN=mean)
stepsPerWeekendInterval <- aggregate(weekendData$allSteps, by=list(interval=weekendData$interval), FUN=mean)
par(mfcol=c(2,1))
plot(stepsPerWeekdayInterval$interval, stepsPerWeekdayInterval$x, type="l", xlab="Interval", ylab="Steps", main="Weekdays")
plot(stepsPerWeekendInterval$interval, stepsPerWeekendInterval$x, type="l", xlab="Interval", ylab="Steps", main="Weekends")

```

It appears that during the weekend the steps count increases.

